---
title: "Visualizing and processing results"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
bibliography: bibliography.bib
---

```{r}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

# `r pretty_region_name(region)`


```{r}
results_directory <- file.path(system.file(paste0("extdata/zeta_results/", region), package = "statisticalRoughness"))
zeta_results <- read_zeta_raster(out_dir = results_directory, Lmax = 1E3, raster_resolution = 10, .len = 10, proxy = FALSE)
rstr <- raster::raster(file.path(system.file("extdata/big_rasters/", package = "statisticalRoughness"), paste0(region, ".tif")))

Lmax <- min(head(dim(rstr), 2))
spatial_scales <- get_all_R_L(Lmax, 5, len = 13)$allL %>% head(-1)
spatial_scales <- spatial_scales[spatial_scales >= 72]
zeta_results$spatial_scales <- tail(spatial_scales, length(zeta_results$raster_list)) * 10
zeta_results$spatial_scales

band_names <- read.csv(file.path(results_directory, "band_names.csv")) %>% names()

if ("theta.x" %in% .selected) zeta_results$raster_list <- zeta_results$raster_list %>% make_angular(match("theta.x", band_names))
if ("theta.y" %in% .selected) zeta_results$raster_list <- zeta_results$raster_list %>% make_angular(match("theta.y", band_names))

```

## DEM

```{r}
rstr <- raster::aggregate(rstr, fact = 10)
library(rayshader)
matrix_for_rayshader(raster::as.matrix(rstr)) %>%
  sphere_shade(colorintensity = 0.5, texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(rstr)), sunaltitude = 20), max_darken = 0.1) %>%  
    # add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(rstr))), 0) %>%
    plot_map()
```

## Maps - clamped - denoised {.tabset}

```{r}
sliced_clamped_raster <- slice_clamp(
  raster_list = zeta_results$raster_list,
  att_names = band_names,
  selected = .selected,
  clamp_raster = TRUE, 
  clamp_params = cparams)
```

```{r}
# sliced_clamped_raster <- sliced_clamped_raster
for (x in c("w.y", "w.x", "xi.y", "xi.x")){
    sliced_clamped_raster <- logtransform_layer(sliced_clamped_raster, match(x, .selected))
}

# sliced_clamped_raster <- reduce_spatial_noise(sliced_clamped_raster)

#' Apply a log transform to a given layer
#' @param raster_list a `list` of `stars` objects
#' @param target_id the id of the layer to be masked
#' @return a `list` of `stars`
#' @export
get_theta_from_theta <- function(raster_list, theta_x, theta_y){
  transformed_raster_list <- lapply(seq_along(raster_list), function(n){
    s <- raster_list[[n]] %>% as("Raster")
    theta_y_values <- raster::getValues(s[[theta_x]]) + 90
    s <- raster::setValues(s, theta_y_values, layer = theta_y)
    return(stars::st_as_stars(s))
  })
  return(transformed_raster_list)
}

sliced_clamped_raster <- get_theta_from_theta(sliced_clamped_raster, match("theta.x", .selected), match("theta.y", .selected))

res1 <- find_par_perp(sliced_clamped_raster, match("alpha1.x", .selected), match("alpha1.y", .selected))
res2 <- find_par_perp(sliced_clamped_raster, match("alpha2.x", .selected), match("alpha2.y", .selected))
sliced_clamped_raster <- res1$raster_list
.selected <- c(.selected, "alpha_perp1", "alpha_par1")

sliced_clamped_raster <- find_par_perp(sliced_clamped_raster, match("alpha1.x", .selected), match("alpha1.y", .selected), match("alpha2.x", .selected), match("alpha2.y", .selected), index_perp = res1$index_perp, index_par = res1$index_par)
.selected <- c(.selected, "alpha_perp2", "alpha_par2")

sliced_clamped_raster <- find_par_perp(sliced_clamped_raster, match("alpha1.x", .selected), match("alpha1.y", .selected), match("w.x", .selected), match("w.y", .selected), index_perp = res1$index_perp, index_par = res1$index_par)
.selected <- c(.selected, "w_perp", "w_par")

sliced_clamped_raster <- find_par_perp(sliced_clamped_raster, match("alpha1.x", .selected), match("alpha1.y", .selected), match("xi.x", .selected), match("xi.y", .selected), index_perp = res1$index_perp, index_par = res1$index_par)
.selected <- c(.selected, "xi_perp", "xi_par")

sliced_clamped_raster <- find_par_perp(sliced_clamped_raster, match("alpha1.x", .selected), match("alpha1.y", .selected), match("theta.x", .selected), match("theta.y", .selected), index_perp = res1$index_perp, index_par = res1$index_par)
.selected <- c(.selected, "theta_perp", "theta_par")

sliced_clamped_raster <- correct_zeta(sliced_clamped_raster, match("zeta1", .selected))
.selected <- c(.selected, "zeta_perp1")
sliced_clamped_raster <- correct_zeta(sliced_clamped_raster, match("zeta2", .selected))
.selected <- c(.selected, "zeta_perp2")


```

```{r, echo = FALSE, results = 'asis'}
res <- lapply(.selected, function(x) {
  knitr::knit_child(text = c(
    '### `r x`',
    '',
    '```{r}',
    'raster_select(sliced_clamped_raster, band_id = match(x, .selected)) %>% 
  make_leaflet_map(ttl = x, groups = zeta_results$spatial_scales, style = "continuous")',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')
```



## Distribution plots {.tabset}

```{r}
ind <- four_values_check(sliced_clamped_raster)
sliced_clamped_raster <- sliced_clamped_raster[ind]
spatial_scales <- zeta_results$spatial_scales[ind]
```

```{r, echo = FALSE, results = 'asis'}
res <- lapply(.selected, function(x) {
  knitr::knit_child(text = c(
    '### `r x`',
    '',
    '```{r}',
    'gb <- make_all_plots(sliced_clamped_raster, spatial_scales, match(x, .selected), x, begin = 0.1, end = 0.85, direction = 1, option = "viridis")',
    'layout_mat <- matrix(c(1, 1, 1, 1, 1, 2, 2, 2, 3, 3), byrow = FALSE, nrow = 5, ncol = 2)',
	'gridExtra::grid.arrange(grobs = gb, layout_matrix = layout_mat)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')
```

## Comparing $x-$ and $y-$ variables

```{r, echo = FALSE, eval = FALSE}
is_band1_gt_band2 <- function(band1, band2, rasters, .selected, spatial_scales){
	allx <- raster_select(rasters, band_id = match(band1, .selected))$rasters
	ally <- raster_select(rasters, band_id = match(band2, .selected))$rasters
	xy <- foreach(x = allx, y = ally, .combine = rbind) %do% {
		x <- as(x, "Raster")
		y <- as(y, "Raster")
		s <- raster::stack(x, y)
		raster::calc(s, fun = function(x){x[1] > x[2]}) %>% raster::getValues() %>% rstatix::freq_table() %>% dplyr::filter(group == "TRUE") %>% dplyr::select(-group)
	}
	# rownames(xy) <- spatial_scales
  mean_xy <- xy %>% dplyr::summarize(n = mean(n), prop = mean(prop))
  # rownames(mean_xy) <- "mean"
  xy <- rbind(xy, mean_xy)
	xy  %>% 
	knitr::kable("html", caption = paste0("Is ", band1, " greater than ", band2, "?")) %>% 
	kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
		position = "center",
		full_width = TRUE)
}
```

```{r, echo = FALSE, results = "asis", eval = FALSE}
vars <- .selected[grepl(".x", .selected)]
res <- foreach(var = vars) %do% {
	is_band1_gt_band2(var, gsub(".x", ".y", var), sliced_clamped_raster, .selected, spatial_scales)
}
cat(unlist(res), sep = '\n')
```

## Cross-scale correlations

```{r, eval = FALSE}
crossscale_correlations(
	raster_list = sliced_clamped_raster,
	selected = .selected,
	spatial_scales = spatial_scales,
	corr_type = "pearson")
```