---
title: "A basic 2D Fourier analysis of topography"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A basic 2D Fourier analysis of topography}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

The code of `fft2d()`, `bin()` and `Hann2D()` is for the most part a port from Matlab to `R` of the code from [2DSpecTools](https://web.mit.edu/perron/www/files/2DSpecTools-v1.1.zip): A 2D spectral analysis toolkit for Matlab by J. Taylor Perron et al.
In consequence, a great introduction to this code is reading @Perron2008 paper in which the authors apply spectral analysis to two examples in California (USA).

# Loading data

First, we load a raster corresponding to one of the two examples from @Perron2008, the one located near Gabilan Mesa in California. 

```{r}
devtools::load_all()
library("statisticalRoughness")
library("rayshader")
library("raster")
gabilan_mesa <- raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "gabilan_mesa.tif"))
```

# Detrending

First, we detrend the DEM using `detrend_dem()` and convert it to a `matrix`.

```{r}
summary(gabilan_mesa)
gabilan_mesa <- gabilan_mesa %>% detrend_dem() 
summary(gabilan_mesa)
```

# Hillshade visualization

Here is what the area looks like.
We can clearly see the imprint of @Perron2008's Fig. 3a.
```{r}
gabilan_mesa %>%
	raster_to_matrix() %>%
	sphere_shade(texture = "desert") %>%
    add_shadow(ray_shade(gabilan_mesa %>% raster_to_matrix(), sunaltitude = 45), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(gabilan_mesa %>% raster_to_matrix()), 0) %>%
    plot_map()
```

# 2D Fourier transform


As `gabilan_mesa` is already a `matrix` that we can pass directly to `fft2d()` with a Hann window (`Hann = TRUE` ensures that `Hann2d()` is called).
We also bin the power spectrum using `bin()` and visualize the results with `spectrum_plot()`.
For convenience, we report as a vertical line the rolloff frequency identified in @Perron2008's Fig. 4a.

```{r}
raster_resolution <- 9.015
FT2D <- fft2D(raster::as.matrix(gabilan_mesa), dx = raster_resolution, dy = raster_resolution, Hann = TRUE)
view_matrix(log10(Re(FT2D$spectral_power_matrix)), ply = FALSE)
```

# Radial power spectrum

While produced with different underlying data, this plot passes the sanity check against @Perron2008's example.

```{r}
nbin <- 50
binned_power_spectrum <- bin(log10(FT2D$radial_frequency_vector), log10(FT2D$spectral_power_vector), nbin)
binned_power_spectrum <- na.omit(binned_power_spectrum)
spectrum_plot(binned_power_spectrum, FT2D) + ggplot2::geom_vline(xintercept = 1/180)
```

# References