---
title: "Computing"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Computing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 7,
  eval = FALSE
)
```
# Description

This script is an example of the use of the function of `statisticalRoughness` on a [tiled DEM](tiling.htlm).


# Initialization

This sets up the `terraOptions` and a reference coordinate system.

```{r}
terra::terraOptions(tempdir = "F:/tmp/")
crs_ref <- terra::crs("+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 + +towgs84=0,0,0,0,0,0,0 +units=m +no_defs ")
```

This sets up the location to look for tiles.

```{r}
dem.dir <- 'data/california-rivers/gis-files/NED'
fname <- "CA_DEM.grd"
fname_sans_ext <- tools::file_path_sans_ext(fname)
out.dir <- file.path(root.dir, dem.dir, paste0(fname_sans_ext, "_tiles"))
```

# Factoring

With this, we retrieve the scales at which the DEM has been tiled.

```{r}
Lmax <- 1E4
allLR <- get_all_R_L(Lmax, 5, only = 12, len = 32)
maxL <- max(allLR$allL)
```

# Computing

Computing can be made parallel or sequential. Here we have 80 tiles and `n` indicates the current tile, `l` indicates the current scale at which the calculation takes place.
Note that the fractal dimension is calculated over the last 5 scales.

```{r}
for (n in seq(80)){
	print(n)
	dem <- terra::rast(file.path(out.dir, paste0(fname_sans_ext, "_tile_", n, ".tif")))
	iters <- floor(seq(from = 1, to = 32))
	for (l in iters){
		tstart <- Sys.time()
		print(l)
		R <- allLR$allRs[[l]] %>% tail(5)
		L <- allLR$allL[[l]]
		fpath <- file.path(out.dir, paste0("Hurst_raster_", n, "_",l , ".tif"))
		if (!file.exists(fpath)){
			Hurst_rasters <- compute_Hurst_rasters(dem, L, R)
			terra::writeRaster(Hurst_rasters, fpath)
		} else {
			Hurst_rasters <- terra::rast(fpath)
		}
		# toc(log = TRUE, quiet = TRUE)
		plot(Hurst_rasters)
		print(Sys.time() - tstart)
	}
	gc()
}
```