---
title: "Factoring"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Factoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, 
  fig.height = 6
)
```

# Introduction

In `statisticalRoughness`, scale factorization is handled by the function `get_all_R_L` which returns a list with the values of $L$, and, for each value of $L$, the associated values of $R$.
The calculation is constrained by a maximum for $L$, `Lmax`, which is _independent_ of map units: it is just a number on the number line.
The second parameter of `get_all_R_L` is the minimum number of factors that have to be found for a value of $L$ to be returned.
Let's see one example:

```{r factor example}
library(statisticalRoughness)
get_all_R_L(20, 3, logfilter = FALSE)
```

There are `r length(get_all_R_L(20, 3, logfilter = FALSE)[[1]])` numbers that have at least 3 factors between 1 and 20.

# `logfilter` option

The `logfilter = FALSE`  option, along with the options detailed below, handles a number of speed up factorization when $L$ becomes large.

`logfilter = TRUE` tries to find $n$ logarithmically spaced factors, with $n$ controlled by the `len` parameter
Let's see it in action with `Lmax = 1E4`:

```{r log example}
get_all_R_L(1e4, 55, logfilter = FALSE, len = 3)
get_all_R_L(1e4, 55, logfilter = TRUE, len = 3)
```

In the first case, `len` has no effect and `r length(get_all_R_L(1e4, 55, logfilter = FALSE, len = 3)[[1]])` are returned.
In the second case, only `r length(get_all_R_L(1e4, 55, logfilter = TRUE, len = 3)[[1]])` logarithmically spaced values are returned.

# `only` option

The `only` options speeds up the execution by looking directly at multiples of 6 and 10 if `only = 610` and at multiples of 12 if `only = 12`.
Because these numbers have the most numbers of factors at the start of the number line, their factors also do have a greater number of factors.
Focusing on these only, speeds up the factorization.
More details can be found [here](fractal_scales.html).

```{r only example}
library(microbenchmark)
tm <- microbenchmark(
	get_all_R_L(1e4, 5, only = NULL),
	get_all_R_L(1e4, 5, only = 610),
	get_all_R_L(1e4, 5, only = 12),
	times = 10
)
ggplot2::autoplot(tm) + ggpubr::theme_pubr()
```