---
title: "Rotating rasters along 2DFT main components"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rotating rasters along 2DFT main components}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

```{r}
devtools::load_all()
```

# Purpose

The main purpose of Fourier analysis in this package is to rotate rasters so that they face the main components of the two-dimensional Fourier spectrum.

# Loading data

First, we perform the basic steps from [this vignette](articles/basic_2dft.html).

```{r}
library("rayshader")
library("raster")
gabilan_mesa <- raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "gabilan_mesa.tif"))
gabilan_mesa <- gabilan_mesa %>% detrend_dem() 
gabilan_mesa <- raster_to_matrix(gabilan_mesa)
raster_resolution <- 9.015
FT2D <- fft2D(gabilan_mesa, dx = raster_resolution, dy = raster_resolution, Hann = TRUE)
nbin <- 50
binned_power_spectrum <- bin(log10(FT2D$radial_frequency_vector), log10(FT2D$spectral_power_vector), nbin)
binned_power_spectrum <- na.omit(binned_power_spectrum)
```

# Filtering the spectral power matrix

Then we obtain a normalized spectral power matrix by removing the background spectrum.
More details are provided in @Perron2008.
This normalized spectral power matrix is then filtered to identify main components.
While @Perron2008 used a comparison with a $\chi^2$ distribution, we filter out the spectral power that are below the 99th percentile of the values.
The plot below corresponds, albeit not exactly, to their Fig. 3c.


```{r}
normalized_spectral_power_matrix <- get_normalized_spectral_power_matrix(binned_power_spectrum, FT2D)
filtered_spectral_power_matrix <- filter_spectral_power_matrix(normalized_spectral_power_matrix, FT2D, quantile_prob = c(0.99))
oce::imagep(filtered_spectral_power_matrix %>% raster_to_matrix(), col = viridisLite::viridis(25), asp = 1)
```

# Extracting the angle

Fromm the filtered spectral power matrix, `get_fourier_angle()` extracts the maximum value of a circular density.
In the current example, that angle $\theta$ is $\simeq$ `r round(get_fourier_angle(filtered_spectral_power_matrix))`.
In their paper, @Perron2008 identifies the main direction at 131$^\circ$ and a secondary one at 45$^\circ$.
The mod(180) we found is `r round(get_fourier_angle(filtered_spectral_power_matrix)) %% 180`.

```{r}
ang_fourier <- get_fourier_angle(filtered_spectral_power_matrix)
ang_fourier
```

```{r, echo = FALSE}
points <- raster::rasterToPoints(filtered_spectral_power_matrix, spatial = TRUE)
coord <- raster::coordinates(points)
coord_polar <- useful::cart2pol(coord[, 1], coord[, 2], degree = TRUE)

power_ww <- na.omit(raster::getValues(filtered_spectral_power_matrix))
power_ww <- log10(power_ww)^2
power_ww <- power_ww / sum(power_ww)
cDens_fourier <- spatstat::circdensity(coord_polar$theta, weights = power_ww, bw = 5)
spatstat::rose(cDens_fourier, main = "Rose plot of FFT components")
```

# Rotating the raster

Now that we have an angle for the rotation, we use the `rotate_raster()` function to turn the raster along the main direction.
This mean that the rows and columns of the underlying matrix are parallel or orthogonal to the main component of the 2D Fourier transform.

```{r}
rotated_raster <- rotate_raster(gabilan_mesa, ang_fourier)
```

## Original topography

```{r}
gabilan_mesa %>%
	sphere_shade(texture = "desert") %>%
    add_shadow(ray_shade(gabilan_mesa, sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(gabilan_mesa), 0) %>%
    plot_map()
```

## Rotated topography

```{r}
rotated_raster %>%
	sphere_shade(texture = "desert") %>%
    add_shadow(ray_shade(rotated_raster, sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(rotated_raster), 0) %>%
    plot_map()
```

```{r}
dim(gabilan_mesa)
dim(rotated_raster)
```

# References