---
title: "Visualizing and processing results"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Visualizing and processing results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

# `r pretty_region_name(.name)`


```{r}
results_directory <- file.path(system.file(paste0("extdata/zeta_results/", .name), package = "statisticalRoughness"))
zeta_results <- read_zeta_raster(out_dir = results_directory, Lmax = 1E3, raster_resolution = 10, .len = 10)
rstr <- raster::raster(file.path(system.file("extdata/big_rasters/", package = "statisticalRoughness"), paste0(.name, ".tif")))

Lmax <- min(head(dim(rstr), 2))
spatial_scales <- get_factors(Lmax) %>% head(-1) %>% tail(length(zeta_results$raster_list)) * 10
zeta_results$spatial_scales <- spatial_scales
zeta_results$spatial_scales

band_names <- read.csv(file.path(results_directory, "band_names.csv")) %>% names()

sliced_clamped_raster <- slice_clamp(
	raster_list = zeta_results$raster_list,
	att_names = band_names,
	selected = .selected,
	clamp_raster = TRUE, 
	clamp_params = cparams)
sliced_clamped_raster %>% head(3)
```

## Maps {.tabset}

```{r, echo = FALSE, results = 'asis'}
res <- lapply(.selected, function(x) {
  knitr::knit_child(text = c(
    '### `r x`',
    '',
    '```{r}',
    'raster_select(sliced_clamped_raster, band_id = match(x, .selected)) %>% 
	make_leaflet_map(ttl = x, groups = zeta_results$spatial_scales, style = "continuous")',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')
```


## Distribution plots {.tabset}

```{r}
ind <- four_values_check(sliced_clamped_raster)
sliced_clamped_raster <- sliced_clamped_raster[ind]
spatial_scales <- zeta_results$spatial_scales[ind]
```

```{r, echo = FALSE, results = 'asis'}
res <- lapply(.selected, function(x) {
  knitr::knit_child(text = c(
    '### `r x`',
    '',
    '```{r}',
    'gb <- make_all_plots(sliced_clamped_raster, spatial_scales, match(x, .selected), x, begin = 0.1, end = 0.85, direction = 1, option = "viridis")',
    'layout_mat <- matrix(c(1, 1, 1, 1, 1, 2, 2, 2, 3, 3), byrow = FALSE, nrow = 5, ncol = 2)',
	'gridExtra::grid.arrange(grobs = gb, layout_matrix = layout_mat)',
    '```',
    ''
  ), envir = environment(), quiet = TRUE)
})
cat(unlist(res), sep = '\n')
```

## Comparing $x-$ and $y-$ variables

```{r, echo = FALSE}
is_band1_gte_band2 <- function(band1, band2, rasters, .selected, spatial_scales){
	allx <- raster_select(rasters, band_id = match(band1, .selected))$rasters
	ally <- raster_select(rasters, band_id = match(band2, .selected))$rasters
	xy <- foreach(x = allx, y = ally, .combine = rbind) %do% {
		x <- as(x, "Raster")
		y <- as(y, "Raster")
		s <- raster::stack(x, y)
		raster::calc(s, fun = function(x){x[1] >= x[2]}) %>% raster::getValues() %>% rstatix::freq_table() %>% dplyr::filter(group == "TRUE") %>% dplyr::select(-group)
	}
	rownames(xy) <- spatial_scales
	xy  %>% 
	knitr::kable("html", caption = paste0("Is ", band1, " greater than or equal to ", band2, "?")) %>% 
	kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
		position = "center",
		full_width = TRUE)
}
```

```{r, echo = FALSE, results = "asis"}
vars <- .selected[grepl(".x", .selected)]
res <- foreach(var = vars) %do% {
	is_band1_gte_band2(var, gsub(".x", ".y", var), sliced_clamped_raster, .selected, spatial_scales)
}
cat(unlist(res), sep = '\n')
```

## Cross-scale correlations

```{r}
crossscale_correlations(
	raster_list = sliced_clamped_raster,
	att_names = band_names,
	selected = .selected,
	spatial_scales = spatial_scales,
	corr_type = "pearson")
```