---
title: "Example of anisotropic landscapes"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example of anisotropic landscapes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 7,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

This vignette displays the calculation of anisotropy exponent for different landscapes: Gabilan Mesa, Yosemite Valley and Modoc Plateau.

```{r, echo = FALSE}
library("statisticalRoughness")
library("rayshader")
view_rotated_raster <- function(rstr, raster_resolution = 9.015, nbin = 50, .Hann = TRUE, .quantile_prob = c(0.9999)){
	rstr <- rstr %>% detrend_dem() 
	FT2D <- fft2D(raster::as.matrix(rstr), dx = raster_resolution, dy = raster_resolution, Hann = .Hann)
	binned_power_spectrum <- bin(log10(FT2D$radial_frequency_vector), log10(FT2D$spectral_power_vector), nbin) %>% na.omit()
	normalized_spectral_power_matrix <- get_normalized_spectral_power_matrix(binned_power_spectrum, FT2D)
	filtered_spectral_power_matrix <- filter_spectral_power_matrix(normalized_spectral_power_matrix, FT2D, quantile_prob = .quantile_prob)
	ang_fourier <- get_fourier_angle(filtered_spectral_power_matrix, FT2D)
	rotated_raster <- rotate_raster(raster::as.matrix(rstr), ang_fourier)
	matrix_for_rayshader(rotated_raster) %>%
		sphere_shade(texture = "desert") %>%
	    add_shadow(ray_shade(matrix_for_rayshader(rotated_raster), sunaltitude = 20), max_darken = 0.3) %>%  
	    add_shadow(ambient_shade(matrix_for_rayshader(rotated_raster)), 0) %>%
	    plot_map()
}
```

# Gabilan Mesa

Gabilan Mesa landscape is the same landscape that we used in other vignettes and the test bed of @Perron2008 methods.
Using `get_zeta()` returns a `data.frame` with the variable of interest for the analysis of topographic roughness and anisotropy.

```{r}
gabilan_mesa <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "gabilan_mesa.tif"))
res_gabi <- get_zeta(gabilan_mesa)
```

Gabilan Mesa landscape has 
a minimum elevation of `r raster::minValue(gabilan_mesa) %>% signif(3)` m,
an average elevation of `r raster::cellStats(gabilan_mesa, stat = "mean") %>% signif(3)` m, 
and a maximum elevation of `r raster::maxValue(gabilan_mesa) %>% signif(3)` m.
The maximum relief is thus `r (raster::maxValue(gabilan_mesa) - raster::minValue(gabilan_mesa)) %>% signif(3)` m.

The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_gabi$theta %% 180 %>% signif(3)`$^\circ$ and
$\theta_y =$ `r (res_gabi$theta %% 180 + 90) %>% signif(3)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_gabi$slope1.beta %>% signif(3)`, and $\beta_2 =$ `r res_gabi$slope2.beta %>% signif(3)`, before and after a cut-off lengthscale of $1/f_c =$ `r (1/res_gabi$change_point.beta) %>% signif(3)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_gabi$slope1.beta) - 2)/2) %>% signif(3)`, and
`r ((abs(res_gabi$slope2.beta) - 2)/2) %>% signif(3)`, respectively.
Before a cut-off lengthscale of
`r mean(c(res_gabi$change_point_mode.x, res_gabi$change_point_mode.y)) %>% signif(3)` m,
the roughness exponents in the $x$-- and $y$--directions are
$\alpha_{1,x} =$ `r res_gabi$slope1_mode.x %>% signif(3)`, and
$\alpha_{1,y} =$ `r res_gabi$slope1_mode.y %>% signif(3)`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_gabi$zeta1 %>% signif(3)`.
After cut-off lengthscale,
the roughness exponents are lower and almost equal with
$\alpha_{2,x} =$ `r res_gabi$slope2_mode.x %>% signif(3)`, and
$\alpha_{2,y} =$ `r res_gabi$slope2_mode.y %>% signif(3)`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_gabi$zeta2 %>% signif(3)`.
The maximum cutoff lengthscale recorded for roughness exponents is $r_{max} =$ `r mean(c(res_gabi$max_point_mode.x, res_gabi$max_point_mode.y)) %>% signif(3)` m.

```{r, echo = FALSE}
matrix_for_rayshader(raster::as.matrix(gabilan_mesa)) %>%
		sphere_shade(texture = "desert") %>%
	    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(gabilan_mesa)), sunaltitude = 20), max_darken = 0.3) %>%  
	    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(gabilan_mesa))), 0) %>%
	    plot_map()
```

# Yosemite Valley

```{r}
yosemite <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "yosemite.tif"))
res_yose <- get_zeta(yosemite) 
```

Yosemite Valley landscape has 
a minimum elevation of `r raster::minValue(yosemite) %>% signif(3)` m,
an average elevation of `r raster::cellStats(yosemite, stat = "mean") %>% signif(3)` m, 
and a maximum elevation of `r raster::maxValue(yosemite) %>% signif(3)` m.
The maximum relief is thus `r (raster::maxValue(yosemite) - raster::minValue(yosemite)) %>% signif(3)` m.
The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_yose$theta %% 180 %>% signif(3)`$^\circ$ and
$\theta_y =$ `r (res_yose$theta %% 180 + 90) %>% signif(3)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_yose$slope1.beta %>% signif(3)`, and $\beta_2 =$ `r res_yose$slope2.beta %>% signif(3)`, before and after a cut-off lengthscale of $1/f_c =$ `r (1/res_yose$change_point.beta) %>% signif(3)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_yose$slope1.beta) - 2)/2) %>% signif(3)`, and
`r ((abs(res_yose$slope2.beta) - 2)/2) %>% signif(3)`.
Before a cut-off lengthscale of
`r mean(c(res_yose$change_point_mode.x, res_yose$change_point_mode.y)) %>% signif(3)` m,
the roughness exponents in the $x$-- and $y$--directions are
$\alpha_{1,x} =$ `r res_yose$slope1_mode.x %>% signif(3)`, and
$\alpha_{1,y} =$ `r res_yose$slope1_mode.y %>% signif(3)`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_yose$zeta1 %>% signif(3)`.
After cut-off lengthscale,
the roughness exponents are lower and almost equal with
$\alpha_{2,x} =$ `r res_yose$slope2_mode.x %>% signif(3)`, and
$\alpha_{2,y} =$ `r res_yose$slope2_mode.y %>% signif(3)`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_yose$zeta2 %>% signif(3)`.
The maximum cutoff lengthscale recorded for roughness exponents is $r_{max} =$ `r mean(c(res_yose$max_point_mode.x, res_yose$max_point_mode.y)) %>% signif(3)` m.


```{r, echo = FALSE}
matrix_for_rayshader(raster::as.matrix(yosemite)) %>%
		sphere_shade(texture = "desert") %>%
	    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(yosemite)), sunaltitude = 20), max_darken = 0.3) %>%  
	    plot_map()
```

# Modoc Plateau

```{r}
modoc <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "modoc.tif"))
res_modo <- get_zeta(modoc) 
```

Modoc Plateau landscape has 
a minimum elevation of `r raster::minValue(modoc) %>% signif(3)` m,
an average elevation of `r raster::cellStats(modoc, stat = "mean") %>% signif(3)` m, 
and a maximum elevation of `r raster::maxValue(modoc) %>% signif(3)` m.
The maximum relief is thus `r (raster::maxValue(modoc) - raster::minValue(modoc)) %>% signif(3)` m.
The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_modo$theta %% 180 %>% signif(3)`$^\circ$ and
$\theta_y =$ `r (res_modo$theta %% 180 + 90) %>% signif(3)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_modo$slope1.beta %>% signif(3)`, and $\beta_2 =$ `r res_modo$slope2.beta %>% signif(3)`, before and after a cut-off lengthscale of $1/f_c =$ `r (1/res_modo$change_point.beta) %>% signif(3)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_modo$slope1.beta) - 2)/2) %>% signif(3)`, and
`r ((abs(res_modo$slope2.beta) - 2)/2) %>% signif(3)`.
Before a cut-off lengthscale of
`r mean(c(res_modo$change_point_mode.x, res_modo$change_point_mode.y)) %>% signif(3)` m,
the roughness exponents in the $x$-- and $y$--directions are almost equal as
$\alpha_{1,x} =$ `r res_modo$slope1_mode.x %>% signif(3)`, and
$\alpha_{1,y} =$ `r res_modo$slope1_mode.y %>% signif(3)`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_modo$zeta1 %>% signif(3)`.
After cut-off lengthscale,
the roughness exponents are more distinct as
$\alpha_{2,x} =$ `r res_modo$slope2_mode.x %>% signif(3)`, and
$\alpha_{2,y} =$ `r res_modo$slope2_mode.y %>% signif(3)`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_modo$zeta2 %>% signif(3)`.
The maximum cutoff lengthscale recorded for roughness exponents is $r_{max} =$ `r mean(c(res_modo$max_point_mode.x, res_modo$max_point_mode.y)) %>% signif(3)` m.


```{r, echo = FALSE}
matrix_for_rayshader(raster::as.matrix(modoc)) %>%
		sphere_shade(texture = "desert") %>%
	    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(modoc)), sunaltitude = 20), max_darken = 0.3) %>%  
	    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(modoc))), 0) %>%
	    plot_map()
```

# Comparing Landscapes

The following table summarizes the metrics for the three landscapes.

| landscape       |               $\beta_1$                |               $\beta_2$                |                     $1/f_c$                      |              $\alpha_{1,x}$              |              $\alpha_{1,y}$              |           $\zeta_{1}$            |              $\alpha_{2,x}$              |              $\alpha_{2,y}$              |            $\zeta_2$             |                                         $r_c$                                         |                                    $r_{max}$                                    |               $\theta_x$                |                   $\theta_y$                   |
|:----------------|:--------------------------------------:|:--------------------------------------:|:------------------------------------------------:|:----------------------------------------:|:----------------------------------------:|:--------------------------------:|:----------------------------------------:|:----------------------------------------:|:--------------------------------:|:-------------------------------------------------------------------------------------:|:-------------------------------------------------------------------------------:|:---------------------------------------:|:----------------------------------------------:|
| Gabilan Mesa    | `r res_gabi$slope1.beta %>% signif(3)` | `r res_gabi$slope2.beta %>% signif(3)` | `r (1/res_gabi$change_point.beta) %>% signif(3)` | `r res_gabi$slope1_mode.x %>% signif(3)` | `r res_gabi$slope1_mode.y %>% signif(3)` | `r res_gabi$zeta1 %>% signif(3)` | `r res_gabi$slope2_mode.x %>% signif(3)` | `r res_gabi$slope2_mode.y %>% signif(3)` | `r res_gabi$zeta2 %>% signif(3)` | `r mean(c(res_gabi$change_point_mode.x, res_gabi$change_point_mode.y)) %>% signif(3)` | `r mean(c(res_gabi$max_point_mode.x, res_gabi$max_point_mode.y)) %>% signif(3)` | `r res_gabi$theta %% 180 %>% signif(3)` | `r (res_gabi$theta %% 180 + 90) %>% signif(3)` |
| Yosemite Valley | `r res_yose$slope1.beta %>% signif(3)` | `r res_yose$slope2.beta %>% signif(3)` | `r (1/res_yose$change_point.beta) %>% signif(3)` | `r res_yose$slope1_mode.x %>% signif(3)` | `r res_yose$slope1_mode.y %>% signif(3)` | `r res_yose$zeta1 %>% signif(3)` | `r res_yose$slope2_mode.x %>% signif(3)` | `r res_yose$slope2_mode.y %>% signif(3)` | `r res_yose$zeta2 %>% signif(3)` | `r mean(c(res_yose$change_point_mode.x, res_yose$change_point_mode.y)) %>% signif(3)` | `r mean(c(res_yose$max_point_mode.x, res_yose$max_point_mode.y)) %>% signif(3)` | `r res_yose$theta %% 180 %>% signif(3)` | `r (res_yose$theta %% 180 + 90) %>% signif(3)` |
| Modoc Plateau   | `r res_modo$slope1.beta %>% signif(3)` | `r res_modo$slope2.beta %>% signif(3)` | `r (1/res_modo$change_point.beta) %>% signif(3)` | `r res_modo$slope1_mode.x %>% signif(3)` | `r res_modo$slope1_mode.y %>% signif(3)` | `r res_modo$zeta1 %>% signif(3)` | `r res_modo$slope2_mode.x %>% signif(3)` | `r res_modo$slope2_mode.y %>% signif(3)` | `r res_modo$zeta2 %>% signif(3)` | `r mean(c(res_modo$change_point_mode.x, res_modo$change_point_mode.y)) %>% signif(3)` | `r mean(c(res_modo$max_point_mode.x, res_modo$max_point_mode.y)) %>% signif(3)` | `r res_modo$theta %% 180 %>% signif(3)` | `r (res_modo$theta %% 180 + 90) %>% signif(3)` |

While the estimate of statistical roughness from $\beta$ and $\alpha$ values differ for the three landscapes, their interpretations are coherent.



# References