---
title: "Example of anisotropic landscapes"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example of anisotropic landscapes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 7,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

This vignette displays the calculation of anisotropy exponent for different landscapes: Gabilan Mesa, Yosemite Valley and Modoc Plateau.

```{r, echo = FALSE}
# library("statisticalRoughness")
devtools::load_all()
library("rayshader")

view_rotated_raster <- function(rstr, raster_resolution = 9.015, nbin = 50, .Hann = TRUE, .quantile_prob = c(0.9999)){
	rstr <- rstr %>% detrend_dem() 
	FT2D <- fft2D(raster::as.matrix(rstr), dx = raster_resolution, dy = raster_resolution, Hann = .Hann)
	binned_power_spectrum <- bin(log10(FT2D$radial_frequency_vector), log10(FT2D$spectral_power_vector), nbin) %>% na.omit()
	normalized_spectral_power_matrix <- get_normalized_spectral_power_matrix(binned_power_spectrum, FT2D)
	filtered_spectral_power_matrix <- filter_spectral_power_matrix(normalized_spectral_power_matrix, FT2D, quantile_prob = .quantile_prob)
	ang_fourier <- get_fourier_angle(filtered_spectral_power_matrix, FT2D)
	rotated_raster <- rotate_raster(raster::as.matrix(rstr), ang_fourier)
	matrix_for_rayshader(rotated_raster) %>%
		sphere_shade(texture = "imhof4") %>%
	    add_shadow(ray_shade(matrix_for_rayshader(rotated_raster), sunaltitude = 20), max_darken = 0.3) %>%  
	    add_shadow(ambient_shade(matrix_for_rayshader(rotated_raster)), 0) %>%
	    plot_map()
}
```

# Gabilan Mesa

Gabilan Mesa landscape is the same landscape that we used in other vignettes and the test bed of @Perron2008 methods.
Using `get_zeta()` returns a `data.frame` with the variable of interest for the analysis of topographic roughness and anisotropy.

```{r}
raster_resolution <- 10
gabilan_mesa <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "gabilan_mesa.tif"))
res_gabi <- get_zeta(gabilan_mesa, raster_resolution) %>% dplyr::mutate_all(signifNA)
```

Gabilan Mesa landscape has 
a minimum elevation of `r raster::minValue(gabilan_mesa)` m,
an average elevation of `r raster::cellStats(gabilan_mesa, stat = "mean")` m, 
and a maximum elevation of `r raster::maxValue(gabilan_mesa)` m.
The maximum relief is thus `r (raster::maxValue(gabilan_mesa) - raster::minValue(gabilan_mesa))` m.

The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_gabi$theta %% 180`$^\circ$ and
$\theta_y =$ `r (res_gabi$theta %% 180 + 90)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_gabi$beta1`, and $\beta_2 =$ `r res_gabi$beta2`, before and after a cut-off lengthscale of $1/f_c =$ `r (res_gabi$inv.fc)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_gabi$beta1) - 2)/2)`, and
`r ((abs(res_gabi$beta2) - 2)/2)`, respectively.
Before a cut-off lengthscale of
`r res_gabi$rc` m,
the roughness exponents in the $x$-- and $y$--directions are
$\alpha_{1,x} =$ `r res_gabi$alpha1.x`, and
$\alpha_{1,y} =$ `r res_gabi$alpha1.y`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_gabi$zeta1`.
After cut-off lengthscale,
the roughness exponents are lower and almost equal with
$\alpha_{2,x} =$ `r res_gabi$alpha2.x`, and
$\alpha_{2,y} =$ `r res_gabi$alpha2.y`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_gabi$zeta2`.

The figure below showcases the landscape in its original form (left) and smoothed with a Gaussian kernel of  `r res_gabi$rc %/% raster_resolution` standard deviations to match the value of $r_c$ (right).

```{r, echo = FALSE}
par(mfrow = c(1,2))
matrix_for_rayshader(raster::as.matrix(gabilan_mesa)) %>%
	sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(gabilan_mesa)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(gabilan_mesa))), 0) %>%
    plot_map()
gabilan_mesa <- gabilan_mesa %>% spatialEco::raster.gaussian.smooth(
		sigma = res_gabi$rc %/% raster_resolution
	)
matrix_for_rayshader(raster::as.matrix(gabilan_mesa)) %>%
	sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(gabilan_mesa)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(gabilan_mesa))), 0) %>%
    plot_map()
```


# Yosemite Valley

```{r}
raster_resolution <- 10
yosemite <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "yosemite.tif"))
res_yose <- get_zeta(yosemite, raster_resolution) %>% dplyr::mutate_all(signifNA)
```

Yosemite Valley landscape has 
a minimum elevation of `r raster::minValue(yosemite)` m,
an average elevation of `r raster::cellStats(yosemite, stat = "mean")` m, 
and a maximum elevation of `r raster::maxValue(yosemite)` m.
The maximum relief is thus `r (raster::maxValue(yosemite) - raster::minValue(yosemite))` m.
The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_yose$theta %% 180`$^\circ$ and
$\theta_y =$ `r (res_yose$theta %% 180 + 90)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_yose$beta1`, and $\beta_2 =$ `r res_yose$beta2`, before and after a cut-off lengthscale of $1/f_c =$ `r (res_yose$inv.fc)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_yose$beta1) - 2)/2)`, and
`r ((abs(res_yose$beta2) - 2)/2)`.
Before a cut-off lengthscale of
`r res_yose$rc` m,
the roughness exponents in the $x$-- and $y$--directions are
$\alpha_{1,x} =$ `r res_yose$alpha1.x`, and
$\alpha_{1,y} =$ `r res_yose$alpha1.y`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_yose$zeta1`.
After cut-off lengthscale,
the roughness exponents are lower and almost equal with
$\alpha_{2,x} =$ `r res_yose$alpha2.x`, and
$\alpha_{2,y} =$ `r res_yose$alpha2.y`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_yose$zeta2`.

The figure below showcases the landscape in its original form (left) and smoothed with a Gaussian kernel of  `r res_yose$rc %/% raster_resolution` standard deviations to match the value of $r_c$ (right).

```{r, echo = FALSE}
par(mfrow = c(1,2))
matrix_for_rayshader(raster::as.matrix(yosemite)) %>%
	sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(yosemite)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(yosemite))), 0) %>%
    plot_map()
yosemite <- yosemite %>% spatialEco::raster.gaussian.smooth(
		sigma = res_yose$rc %/% raster_resolution
	)
matrix_for_rayshader(raster::as.matrix(yosemite)) %>%
	sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(yosemite)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(yosemite))), 0) %>%
    plot_map()
```


# Modoc Plateau

```{r}
raster_resolution <- 10
modoc <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "modoc.tif"))
res_modo <- get_zeta(modoc, raster_resolution) %>% dplyr::mutate_all(signifNA)
```

Modoc Plateau landscape has 
a minimum elevation of `r raster::minValue(modoc)` m,
an average elevation of `r raster::cellStats(modoc, stat = "mean")` m, 
and a maximum elevation of `r raster::maxValue(modoc)` m.
The maximum relief is thus `r (raster::maxValue(modoc) - raster::minValue(modoc))` m.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_modo$theta %% 180`$^\circ$ and
$\theta_y =$ `r (res_modo$theta %% 180 + 90)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_modo$beta1`, and $\beta_2 =$ `r res_modo$beta2`, before and after a cut-off lengthscale of $1/f_c =$ `r (res_modo$inv.fc)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_modo$beta1) - 2)/2)`, and
`r ((abs(res_modo$beta2) - 2)/2)`.
Before a cut-off lengthscale of
`r res_modo$rc` m,
the roughness exponents in the $x$-- and $y$--directions are almost equal as
$\alpha_{1,x} =$ `r res_modo$alpha1.x`, and
$\alpha_{1,y} =$ `r res_modo$alpha1.y`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_modo$zeta1`.
After cut-off lengthscale,
the roughness exponents are more distinct as
$\alpha_{2,x} =$ `r res_modo$alpha2.x`, and
$\alpha_{2,y} =$ `r res_modo$alpha2.y`.
However, the anisotropy exponent, $\zeta_2$, cannot be derived as at least one ofthe distribution of $\alpha_2$ has a large spread and its mean is an inconclusive description of centrality.

The figure below showcases the landscape in its original form (left) and smoothed with a Gaussian kernel of  `r res_modo$rc %/% raster_resolution` standard deviations to match the value of $r_c$ (right).


```{r, echo = FALSE}
par(mfrow = c(1,2))
matrix_for_rayshader(raster::as.matrix(modoc)) %>%
	sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(modoc)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(modoc))), 0) %>%
    plot_map()
modoc <- modoc %>% spatialEco::raster.gaussian.smooth(
		sigma = res_modo$rc %/% raster_resolution
	)
matrix_for_rayshader(raster::as.matrix(modoc)) %>%
	sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(modoc)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(modoc))), 0) %>%
    plot_map()
```

# Marble Canyon



```{r}
raster_resolution <- 90
marble_canyon <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "marble_canyon.tif"))
marble_canyon <- raster::aggregate(marble_canyon, fact = 9) # 90 m to match Pastor-Satorras
res_mrbl <- get_zeta(marble_canyon, raster_resolution) %>% dplyr::mutate_all(signifNA)
```

Marble Canyon landscape has 
a minimum elevation of `r raster::minValue(marble_canyon)` m,
an average elevation of `r raster::cellStats(marble_canyon, stat = "mean")` m, 
and a maximum elevation of `r raster::maxValue(marble_canyon)` m.
The maximum relief is thus `r (raster::maxValue(marble_canyon) - raster::minValue(marble_canyon))` m.

The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_mrbl$theta %% 180`$^\circ$ and
$\theta_y =$ `r (res_mrbl$theta %% 180 + 90)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_mrbl$beta1`, and $\beta_2 =$ `r res_mrbl$beta2`, before and after a cut-off lengthscale of $1/f_c =$ `r (res_mrbl$inv.fc)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_mrbl$beta1) - 2)/2)`, and
`r ((abs(res_mrbl$beta2) - 2)/2)`, respectively.
Before a cut-off lengthscale of
`r res_mrbl$rc` m,
the roughness exponents in the $x$-- and $y$--directions are
$\alpha_{1,x} =$ `r res_mrbl$alpha1.x`, and
$\alpha_{1,y} =$ `r res_mrbl$alpha1.y`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_mrbl$zeta1`.
After cut-off lengthscale,
the roughness exponents are lower and almost equal with
$\alpha_{2,x} =$ `r res_mrbl$alpha2.x`, and
$\alpha_{2,y} =$ `r res_mrbl$alpha2.y`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_mrbl$zeta2`.

The figure below showcases the landscape in its original form (left) and smoothed with a Gaussian kernel of  `r res_mrbl$rc %/% raster_resolution` standard deviations to match the value of $r_c$ (right).

```{r, echo = FALSE}
par(mfrow = c(1,2))
matrix_for_rayshader(raster::as.matrix(marble_canyon)) %>%
  sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(marble_canyon)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(marble_canyon))), 0) %>%
    plot_map()
marble_canyon <- marble_canyon %>% spatialEco::raster.gaussian.smooth(
    sigma = res_gabi$rc %/% raster_resolution
  )
matrix_for_rayshader(raster::as.matrix(marble_canyon)) %>%
  sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(marble_canyon)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(marble_canyon))), 0) %>%
    plot_map()
```

# Submarine Canyon

```{r}
raster_resolution <- 50
submarine_canyon <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "submarine_canyon.tif"))
res_subm <- get_zeta(submarine_canyon, raster_resolution) %>% dplyr::mutate_all(signifNA)
```

Submarine Canyon landscape has 
a minimum elevation of `r raster::minValue(submarine_canyon)` m,
an average elevation of `r raster::cellStats(submarine_canyon, stat = "mean")` m, 
and a maximum elevation of `r raster::maxValue(submarine_canyon)` m.
The maximum relief is thus `r (raster::maxValue(submarine_canyon) - raster::minValue(submarine_canyon))` m.

The figure below showcases the landscape in its original form.
The landscape was rotated so that in this original frame of reference,
$\theta_x =$ `r res_subm$theta %% 180`$^\circ$ and
$\theta_y =$ `r (res_subm$theta %% 180 + 90)`$^\circ$.
(In the rotated frame of reference, $\theta_x' = 0^\circ$ and $\theta_y' = 90^\circ$.)
The segmented regression on the power spectrum identifies scalings with slope $\beta_1 =$ `r res_subm$beta1`, and $\beta_2 =$ `r res_subm$beta2`, before and after a cut-off lengthscale of $1/f_c =$ `r (res_subm$inv.fc)` m, respectively.
As $\beta = 2H + d$, and $d = 2$, the Hurst coefficients $H$ corresponding to $\beta_1$ and $\beta_2$ are 
`r ((abs(res_subm$beta1) - 2)/2)`, and
`r ((abs(res_subm$beta2) - 2)/2)`, respectively.
Before a cut-off lengthscale of
`r res_subm$rc` m,
the roughness exponents in the $x$-- and $y$--directions are
$\alpha_{1,x} =$ `r res_subm$alpha1.x`, and
$\alpha_{1,y} =$ `r res_subm$alpha1.y`.
Consequently, the anisotropy exponent is $\zeta_1 =$ `r res_subm$zeta1`.
After cut-off lengthscale,
the roughness exponents are lower and almost equal with
$\alpha_{2,x} =$ `r res_subm$alpha2.x`, and
$\alpha_{2,y} =$ `r res_subm$alpha2.y`.
Consequently, the anisotropy exponent is $\zeta_2 =$ `r res_subm$zeta2`.

The figure below showcases the landscape in its original form (left) and smoothed with a Gaussian kernel of  `r res_subm$rc %/% raster_resolution` standard deviations to match the value of $r_c$ (right).

```{r, echo = FALSE}
par(mfrow = c(1,2))
matrix_for_rayshader(raster::as.matrix(submarine_canyon)) %>%
  sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(submarine_canyon)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(submarine_canyon))), 0) %>%
    plot_map()
submarine_canyon <- submarine_canyon %>% spatialEco::raster.gaussian.smooth(
    sigma = res_gabi$rc %/% raster_resolution
  )
matrix_for_rayshader(raster::as.matrix(submarine_canyon)) %>%
  sphere_shade(texture = "imhof4") %>%
    add_shadow(ray_shade(matrix_for_rayshader(raster::as.matrix(submarine_canyon)), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(raster::as.matrix(submarine_canyon))), 0) %>%
    plot_map()
```

# Comparing Landscapes

The following table summarizes the metrics for the three landscapes.

```{r}
res <- rbind(res_gabi, res_yose, res_modo, res_mrbl, res_subm) %>% 
  dplyr::select(-c("w", "xi", "alpha1", "alpha2"))
rownames(res) <- c("Gabilan Mesa", "Yosemite Valley", "Modoc Plateau", "Marble Canyon", "Submarine Canyon")
```

```{r, echo = FALSE}
cnames <- c(
  "$\\beta_1$",
  "$\\beta_2$",
  "$\\alpha_{1,x}$",
  "$\\alpha_{1,y}$",
  "$\\zeta_1$",
  "$\\alpha_{2,x}$",
  "$\\alpha_{2,y}$",
  "$\\zeta_2$",
  "$\\theta$",
  "$1/f_c$",
  "$r_c$",
  "$\\xi_x$",
  "$\\xi_y$",
  "$w_x$",
  "$w_y$"
)
res %>% knitr::kable("html", col.names = cnames) %>% 
kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),
  position = "center",
  full_width = FALSE) %>% 
kableExtra::scroll_box(width = "7in", height = "3in") 
```


While the estimate of statistical roughness from $\beta$ and $\alpha$ values differ for the three landscapes, their interpretations are coherent.



Gabilan Mesa landscape is weakly anisotropic and mainly correlated below `r res_gabi$rc` m, and isotropic and mainly anti-correlated above `r res_gabi$rc` m.
Yosemite Valley landscape is anisotropic and mainly correlated below `r res_yose$rc` m, and isotropic with no clear correlation above `r res_yose$rc` m.
Modoc Plateau landscape is isotropic and mainly correlated below `r res_modo$rc` m, and with undetermined behaviour above `r res_modo$rc` m.



# References