---
title: "Deriving the anisotropy exponent"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deriving the anisotropy exponent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

This vignette showcases the functions used to derive the roughness and anisotropy exponents from a rotated landscape.

# Loading data

First, we perform all the steps from [this vignette](articles/rotating_rasters.html).

```{r loading}
library("statisticalRoughness")
```

First, we load one of the DEM.

```{r reading rasters}
gabilan_mesa <- terra::rast(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "gabilan_mesa.tif"))
```

Next we create tiles over which we will calculate the roughness and anisotropy exponents.
This is done for simplicity here by using `terra::aggregate()`.

```{r tiles}
tiles <- terra::aggregate(gabilan_mesa, fact = 100)
```

Then, `get_zeta_df()` basically checks, conversions and parallelization to be able to efficiently run `get_zeta()`.
It is also possible to directly to pass a polygon to `get_zeta_df()`.
If passing a raster, the conversion from raster to polygons is handled internally.

```{r get_zeta_df}
zeta_df <- get_zeta_df(
  gabilan_mesa, tiles, 
  crs_ref = terra::crs("EPSG:3310"), vertical_accuracy = 1.87
)
```

`get_zeta_df()` returns a `data.frame` with no spatial information.
`get_zeta_raster()` takes care to re-introduce the spatial information from the initial raster (here, `gabilan_mesa`) and from the `tiles`.
If the `.zeta_df` argument is set to `NULL`, `get_zeta_raster()` first runs `get_zeta()`

```{r get_zeta_raster}
zeta_raster <- get_zeta_raster(
  gabilan_mesa, tiles, 
  .zeta_df = zeta_df, crs_ref = terra::crs("EPSG:3310"), vertical_accuracy = 1.87
)
names(zeta_raster)
```

