---
title: "Running the workflow for tiles"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the workflow for tiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

This vignette showcases the functions used to derive the roughness and anisotropy exponents for a collection of tiles over a landscape.

# Loading data

First, we perform all the steps from [this vignette](articles/rotating_rasters.html).

```{r loading}
library("statisticalRoughness")
```

First, we load one of the DEM.

```{r reading rasters}
rstr <- raster::raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "gabilan_mesa.tif"))
raster_resolution <- 10
```

Next we create tiles over which we will calculate the roughness and anisotropy exponents.
This is done for simplicity here by using `raster::aggregate()`.

```{r tiles}
tiles <- raster::aggregate(rstr, fact = 100)
```

# Tiled workflow

Then, `get_zeta_df()` basically checks, conversions and parallelization to be able to efficiently run `get_zeta()`.
It is also possible to directly to pass a polygon to `get_zeta_df()`.
If passing a raster, the conversion from raster to polygons is handled internally.

```{r get_zeta_df}
zeta_df <- get_zeta_df(
  rstr, tiles, 
  raster_resolution, vertical_accuracy = 1.87
)
```

`get_zeta_df()` returns a `data.frame` with no spatial information.
`get_zeta_raster()` takes care to re-introduce the spatial information from the initial raster (here, `rstr`) and from the `tiles`.
If the `.zeta_df` argument is set to `NULL`, `get_zeta_raster()` first runs `get_zeta()`

```{r get_zeta_raster}
zeta_raster <- get_zeta_raster(
  rstr, tiles, 
  .zeta_df = zeta_df, raster_resolution, vertical_accuracy = 1.87
)
```

For example here are the raster layers corresponding to $\alpha_{1,x}$ and $\alpha_{1,y}$.

```{r visualization, echo = FALSE}
sp::spplot(zeta_raster[["alpha1.x"]])
sp::spplot(zeta_raster[["alpha1.y"]])
```