---
title: "Deriving the anisotropy exponent"
author:
- name: Hervé Guillon
date: "`r format(Sys.time(), '%d %B %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Deriving the anisotropy exponent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 8,
  warning = FALSE,
  message = FALSE
)
```

# Purpose

This vignette showcases the functions used to derive the roughness and anisotropy exponents from a rotated landscape.

# Loading data

First, we perform all the steps from [this vignette](articles/rotating_rasters.html).

```{r loading}
library("statisticalRoughness")
library("rayshader")
library("raster")

gabilan_mesa <- raster(file.path(system.file("extdata/rasters/", package = "statisticalRoughness"), "modoc.tif"))
gabilan_mesa <- gabilan_mesa %>% detrend_dem() 
raster_resolution <- 9.015
FT2D <- fft2D(raster::as.matrix(gabilan_mesa), dx = raster_resolution, dy = raster_resolution, Hann = TRUE)
nbin <- 20
binned_power_spectrum <- bin(log10(FT2D$radial_frequency_vector), log10(FT2D$spectral_power_vector), nbin)
binned_power_spectrum <- na.omit(binned_power_spectrum)
normalized_spectral_power_matrix <- get_normalized_spectral_power_matrix(binned_power_spectrum, FT2D)
filtered_spectral_power_matrix <- filter_spectral_power_matrix(normalized_spectral_power_matrix, FT2D, quantile_prob = c(0.9999))
ang_fourier <- get_fourier_angle(filtered_spectral_power_matrix, FT2D)
rotated_raster <- rotate_raster(raster::as.matrix(gabilan_mesa), ang_fourier)
```

On the figure below, the horizontal line is the middle row, the leftmost vertical line is the first third column and the rightmost vertical line is the middle column.

```{r rayshader, echo = FALSE}
ind_x <- nrow(rotated_raster) %/% 2
ind_y <- ncol(rotated_raster) %/% 3
.rotated_raster <- rotated_raster
.rotated_raster[ind_x, ] <- NA
.rotated_raster[, ind_y] <- NA
.rotated_raster[, ind_x] <- NA

matrix_for_rayshader(.rotated_raster) %>%
	sphere_shade(texture = "desert") %>%
    add_shadow(ray_shade(matrix_for_rayshader(.rotated_raster), sunaltitude = 20), max_darken = 0.3) %>%  
    add_shadow(ambient_shade(matrix_for_rayshader(.rotated_raster)), 0) %>%
    plot_map()
```

# Height-height correlation functions

The height-height correlation function (HHCF) is defined as:

\[C(r) = \left\langle \sqrt{[h(x+r,t)-h(x,t)]^2} \right\rangle\]

with $h$ the elevation, $x$ the spatial coordinate, $t$ the time coordinate, $r$ a spatial increment and where brackets indicate averaging.
With this definition, we can now calculate the HHCF for the `rotated_raster` in the row-wise (`margin = 1`) and column-wise directions (`margin = 2`) using `get_hhcf()`.

```{r hhcf}
hhcf_x <- get_hhcf(rotated_raster, margin = 1)
hhcf_y <- get_hhcf(rotated_raster, margin = 2)
```

# Roughness exponent, $\alpha$

## Definition and examples

The HHCF scales with increasing $r$ as: 

\[ C(r) \sim r^\alpha \]

with $\alpha$ the roughness exponent.
We now use `get_alpha()` to derive the roughness exponents $\alpha_x$ and $\alpha_y$ in each direction.
The option `do_plot = TRUE` provides a graphical display of the step-wise binned regression process and identification of the cross-over lengthscale.
As the regression is a piecewise regression, for each direction, two slopes are returned, for example: $\alpha_{1,x}$ and $\alpha_{2,x}$.

```{r example_x_1}
get_alpha(hhcf_x$hhcf[ind_x, ], raster_resolution, do_plot = TRUE)
```

```{r example_y_1}
get_alpha(hhcf_y$hhcf[ind_y, ], raster_resolution, do_plot = TRUE)
```

`get_alpha()` includes a number of safeguard and the function will return `NA`.
Here is an example the roughly follow a ridgeline, the middle column, with no statistically consistent scaling.

```{r example_y_2}
get_alpha(hhcf_y$hhcf[ind_x, ], raster_resolution, do_plot = TRUE)
```

## Computing all roughness exponents

Provided a list of HHCF from `get_hhcf()`, the function `get_all_alpha()` computes all step-wise regressions.

```{r all_alpha}
alpha_x <- get_all_alpha(hhcf_x, raster_resolution) %>% filter_alpha() 
alpha_y <- get_all_alpha(hhcf_y, raster_resolution) %>% filter_alpha()
```

```{r, echo = FALSE}
alphas <- rbind(
  alpha_x %>% stats::na.omit() %>% dplyr::mutate(direction = 'x'),
  alpha_y %>% stats::na.omit() %>% dplyr::mutate(direction = 'y')
  ) %>% reshape2::melt(measure.vars = c("slope1", "slope2"))
p <- ggplot2::ggplot(alphas, ggplot2::aes(x = .data$value, group = .data$direction, color = .data$direction)) +
  ggplot2::geom_density(n = 512) +
  ggplot2::facet_grid(.data$variable ~ .) +
  ggpubr::theme_pubr()
p
```


The resulting `data.frames` can easily be summarized with `summarise_alpha()`.
For example:

```{r summarize}
alpha_x %>% summarise_alpha()
alpha_y %>% summarise_alpha()
```

```{r summarize_hidden, echo = FALSE}
x_stats <- alpha_x %>% summarise_alpha()
y_stats <- alpha_y %>% summarise_alpha()
```

In a more human-readable format, we can summarize the entire landscape on `rotated_raster` by:

|            |                             row-wise direction, $x$                             |                           column-wise direction, $y$                            |
|:-----------|:-------------------------------------------------------------------------------:|:-------------------------------------------------------------------------------:|
| $\alpha_1$ | `r x_stats$slope1_mode %>% signif(2)` $\pm$ `r x_stats$slope1_sd %>% signif(2)` | `r y_stats$slope1_mode %>% signif(2)` $\pm$ `r y_stats$slope1_sd %>% signif(2)` |
| $\alpha_2$ | `r x_stats$slope2_mode %>% signif(2)` $\pm$ `r x_stats$slope2_sd %>% signif(2)` | `r y_stats$slope2_mode %>% signif(2)` $\pm$ `r y_stats$slope2_sd %>% signif(2)` |

These numbers mean that the topography is mainly correlated in a positive way below `r mean(c(x_stats$change_point_mode, y_stats$change_point_mode)) %>% signif(3)` meters and more correlated in the $x$-direction than in the $y$-direction as $\alpha_{1,x} > \alpha_{1,y} > 0.5$.
Above `r mean(c(x_stats$change_point_mode, y_stats$change_point_mode)) %>% signif(3)` meters, the topography is negatively correlated and more so in the y-direction than in the $x$-direction $\alpha_{2,y} < \alpha_{2,x} < 0.5$.
These results also mean that the $x$-direction corresponds to the down-slope direction, $\mathbf{e}_\parallel$, while the $y$-direction corresponds to the across-slope direction, $\mathbf{e}_\perp$.
Further, the dominant slope in the $x$-direction results in increased smoothing and higher values of roughness exponents.
This is what we could have expected from a visual inspection of the landscape.

# Anisotropy exponent, $\zeta$

The anisotropy exponent is the ratio between the down-slope and across-slope roughness exponents:

\[ \zeta = \frac{\alpha_\perp}{\alpha_\parallel}\]

Because of the increased smoothing due to gravity in the down-slope direction, $\alpha_\perp \geq \alpha_\parallel$ and $\zeta \geq 1$.
From the central values of $\alpha_{1,x}$ and $\alpha_{1,y}$, the value of the anisotropy exponent $\zeta_1$ can be derived as:

\[ \zeta_1 = \frac{\max{\alpha_{1,x}, \alpha_{1,x}}}{\min{\alpha_{1,x}, \alpha_{1,x}}}\]

This estimate of $\zeta$ is handled by an internal function. For the current landscape:

|           |                     anisotropy exponent, $\zeta$                      |
|:----------|:---------------------------------------------------------------------:|
| $\zeta_1$ | `r get_zeta_(x_stats$slope1_mode, y_stats$slope1_mode) %>% signif(3)` |
| $\zeta_2$ | `r get_zeta_(x_stats$slope2_mode, y_stats$slope2_mode) %>% signif(3)` |
